>

### 2020-05-13 manuscript/manuscript.Rmd
Rebuild table 2 with xtable
This helps make the table more readable, with
columns that align with each other. Some other edits to tables in the primary paper body. Addresses #7

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
library(qwraps2)
library(knitr)
library(kableExtra)
library(texreg)
library(here)

@@

```{r tracts}
# built in 01_health_data.Rmd and 03_logsums.Rmd
tracts <- read_rds(here("data/tracts_access.rds"))
```

@@

For comparative purposes, we also employ an isochrone analysis using the
10-minute walk threshold calculated for ParkScore [@parkscore2019]. If the
population-weighted centroid of a tract is within a 10-minute walk of a green space
as defined by ParkScore, the tract is considered as having "access" to a park.
The centroid is necessary as all tracts in New York City have at least *some* 
intersection with the 10-minute walk buffer; as it is, only 
`r sum(!tracts$walk_10)` of the `r nrow(tracts)` tract centroids are not located 
within this buffer.

```{r map, fig.env="figure*", echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.cap="Normalized utility-based park accessibility values in New York City."}
library(ggmap)
map <- ggmap(
  get_stamenmap(unname(st_bbox(tracts)), 
                maptype = "toner-lite"), extent = "device")
map +
  geom_sf(data = tracts, inherit.aes=FALSE,
          aes(fill = cut(multi_ls, breaks = c(-Inf, -3, -2, -1, 0, 1, 2, 3, Inf))),
          alpha = 0.9, color = NA) +
  scale_fill_brewer("Accessibility", palette = "RdYlBu")
```

@@

```{r weights, warning = FALSE}
# This are developed in 03_logsums.Rmd
distances <- read_rds(here("data/distances.rds"))
W <- read_rds(here("data/W.rds"))
trMC <- trW(as(W, "CsparseMatrix"), type="MC")
```

@@

names_for_terms <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1])
)

lambda <- "$\\lambda$: spatial correlation"
access_terms <- c("Logsum", "Tweets",  "Attributes", "10-min walk")
```

@@

To determine which spatial effects specification was appropriate, we estimated
SLX and SDEM models of physical activity participation rates as a function of
tract covariates and  accessibility. These two models are presented in Table 
\ref{tab:slx-sdem}. With only a few exceptions, the SLX and SDEM coefficient
estimates for corresponding covariates lie within the 95\% confidence intervals
of the other model. A Hausman-style test [@Pace2008] on the coefficients 
assumes a null hypothesis that the coefficients are equivalent; this test 
produces $p$-value of only $`r slx_sdem_htest[2]`$. Though this statistic is
technically beneath a common hypothesis test threshold of $\alpha = 0.05$, we
fail to reject the null hypothesis that the coefficients are equivalent, as 
the differences are at no point practically different from each other. The consequences
of rejecting the null hypothesis in this case would be to suggest a model with 
spatial dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we adopt the
SDEM specification moving forward as the spatial lag of the errors is 
significant and the overal log likelihood of the SDEM is higher than the SLX 
specification.

\begin{table*}
\caption{Comparison of SLX and SDEM Coefficients}
\label{tab:slx-sdem}
```{r slx-sdem, results = "asis"}
texreg(slx_vs_sdem, single.row = TRUE, ci.force = TRUE,  digits = 4,
       custom.coef.names = c(names_for_terms, "Accessibility", lambda), 
       # Goodness of fit (LM and SP are different)
       custom.gof.rows = list(
         "Log Likelihood" = lapply(slx_vs_sdem, logLik) %>% unlist(),
         "Num. Obs." = lapply(slx_vs_sdem, function(s) length(s$residuals))
       ),
       include.rsquared = F, include.adjrs = F, include.aic = F, include.loglik = F,
       include.parameters = F, include.nobs = F,
       use.packages = FALSE, booktabs = TRUE,
       caption = "Comparison of SLX and SDEM Coefficients",
       label = "tab:slx-sdem",table = FALSE,
       custom.note = "%stars $y = $ tract-level physical activity rate.",
       caption.above = TRUE)
```
\end{table*}

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and accessibility
coefficients for three models using different specifications of the utility
equation, as well as the ten-minute walk buffer calculated by ParkScore. 
Note that the model called "Size and Distance" is precisely the model
labeled "SDEM" in Table \ref{tab:slx-sdem}. The controlling covariates are
suppressed for clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in the Appendix. Each of the three
utility-based specifications is significant at the $\alpha = 0.05$ level or lower,
and is of the hypothesized directionality. That is, residents living in tracts 
with increased utility-based accessiblity to green spaces show a significantly
higher physical activity participation rate. Taking the rough mean of the 2
estimated utility-based coefficients, moving from the least-accessible tract
($A_i = -3$) to the
most-accessible tract ($A_i = 3$) is expected to raise the physical activity rate
by `r 6 * .15` percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor park access
is anywhere between `r 6 * .02` and `r 6 * .3` percentage points.

@@

\begin{table*}
\caption{Estimated Effect of Accessibility on Physical Activity Rates}
\label{tab:pa-models}
```{r pa-models, results = "asis"}
texreg(pa_models[c(2, 4:5)], digits = 4, ci.force = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "lambda" = "$\\lambda$: spatial correlation", 
         "access_ls" = "Size and Distance", 
         "multi_ls" = "Amenities", 
         "walk_10TRUE" = "10-minute walk"),
       caption = "Estimated Effect of Accessibility on Physical Activity Rates",
       label = "tab:pa-models", table = FALSE,
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```
\end{table*}

@@

As far as the accessibility values are concerned, none of the three
utility-based specifications nor the ParkScore 10-minute walk buffer are
correlated with obesity rates at the traditional 95\% confidence interval. 
It is important to consider the implications of the full range of the confidence
interval, however [@Amrhein2019]. The results suggest that after controlling
for tract socioeconomic variables, physical activity rates, and spatially correlated
error terms, obesity rates in tracts with excellent park accessibility are expected 
to be between `r 6 * 0.15` percentage points lower and `r 6 * .0053` percentage 
points higher than tracts with poor park accessibility. This range of values 
suggests that accessibility to parks may have obesity-reducing effects beyond
the effect of accessibility on physical activity rates worth further study.

\begin{table*}
\caption{Estimated Effect of Accessibility on Obesity Rates}
\label{tab:obesity-models}
```{r obesity-models, results="asis"}
texreg(obesity_models[c(2, 4:5)], digits = 4, ci.force = TRUE, 
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "physact" = "Physical Activity", 
         "lag.physact" = "$\\gamma$: Physical Activity",
         "lambda" = "$\\lambda$: spatial correlation", 
         "access_ls" = "Size and Distance", 
         "multi_ls" = "Amenities", 
         "walk_10TRUE" = "10-minute walk"),
       caption = "Estimated Effect of Accessibility on Obesity Rates",
       label = "tab:obesity-models", table = FALSE,
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```
\end{table*}

@@

Finally, this study is primarily focused on the hypothesis that accessibility to
parks encourages physical activity, which in turn reduces obesity. There are a
multitude of other hypotheses that might be proposed and tested with the basic
methodology we have employed here, or competing explanations for the outcomes we
have observed. It is distinctly possible, for instance, that individuals who wish
to exercise regularly in parks choose to live near them. Given that the CDC models
informing the small-area obesity and physical activity estimates presumably include variables
likely to influence such preferences (income, etc.), our investigation cannot
isolate the preferences from the effect. Controlling for such a
self-selection effect would be necessary to isolate the exogenous
impacts of park access on obesity or other health outcomes. And regarding these
other variables: this study did not consider potential relationships between park
access and hospitalization rates, life expectancy, respiratory disease,
mental health, or any number of potential beneficial outcomes hypothesized
or explored in the existing literature. Exploring these connections and their
underlying mechanisms should be a priority as city planners and urban architects
attempt to improve the quality of life of urban residents in the future.

@@

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
accessibility and these metrics, previous literature has produced mixed
findings, perhaps owing to the range of variables modeled and the coarse spatial
scale of the analyses. Using New York City as a case study, we presented a
holistic and flexible measurement for park accessibility that compares the
continuous distance to all parks in the region, weighted against the size of the
park and its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive relationship
where the least park-accessible tracts have an expected physical activity
participation rate between `r 6 * .05` and `r 6 * .35` percentage points lower than the most
accessible tracts. We also found a suggestive though not sigificant relationship
between park accessiblity and expected obesity rates *in addition to* the effect
of physical activity participation.

>>>>

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
library(qwraps2)
library(knitr)
library(kableExtra)
library(texreg)
library(here)
library(xtable)

@@

```{r tracts}
# built in 01_health_data.Rmd and 04_accessibility.Rmd
tracts <- est_data <- read_rds(here("data/tracts.rds")) %>%
  left_join(read_rds(here("data/tract_logsums.rds")), by = "geoid") %>%
  left_join(read_rds(here("data/walk10.rds")), by = "geoid")
```

@@

For comparative purposes, we also employ an isochrone analysis using the
10-minute walk threshold calculated for ParkScore [@parkscore2019]. If the
population-weighted centroid of a tract is within a 10-minute walk of a green space
as defined by ParkScore, the tract is considered as having "access" to a park.
The centroid is necessary as all tracts in New York City have at least *some* 
intersection with the 10-minute walk buffer; as it is, only 
`r sum(!tracts$walk10)` of the `r nrow(tracts)` tract centroids are not located 
within this buffer.

```{r map, fig.env="figure*", echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.cap="Normalized utility-based park accessibility values in New York City."}
library(ggmap)
map <- ggmap(
  get_stamenmap(unname(st_bbox(tracts)), 
                maptype = "toner-lite"), extent = "device")
map +
  geom_sf(data = tracts, inherit.aes=FALSE,
          aes(fill = cut(euc_multi, breaks = c(-Inf, -3, -2, -1, 0, 1, 2, 3, Inf))),
          alpha = 0.9, color = NA) +
  scale_fill_brewer("Accessibility", palette = "RdYlBu")
```

@@

```{r weights, warning = FALSE}
W <- read_rds(here("data/Wtr.rds"))
trMC <- trW(as(W, "CsparseMatrix"), type="MC")
```

@@

termnames <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1]), 
  "Accessibility", "$\\lambda$: spatial correlation"
)

#coefficient map
cm <- as.list(termnames)
names(cm) <- c(names(slx_vs_sdem$SDEM$coefficients), "$\\lambda$")
```

@@

To determine which spatial effects specification was appropriate, we estimated
SLX and SDEM models of physical activity participation rates as a function of
tract covariates and  accessibility. These two models are presented in Table 
\ref{tab:slx-sdem}. With only a few exceptions, the SLX and SDEM coefficient
estimates for corresponding covariates lie within the 95\% confidence intervals
of the other model. A Hausman-style test [@Pace2008] on the coefficients 
assumes a null hypothesis that the coefficients are equivalent; this test 
produces $p$-value of only $`r slx_sdem_htest[2]`$. Though this statistic is
well beneath a common hypothesis test threshold of $\alpha = 0.05$, we
elect retain the null hypothesis that the coefficients are equivalent, as 
the differences are at no point practically different from each other. The consequences
of rejecting the null hypothesis in this case would be to suggest a model with 
spatial dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we adopt the
SDEM specification moving forward as the spatial lag of the errors is 
significant and the overal log likelihood of the SDEM is higher than the SLX 
specification.

\begin{table*}
\caption{Comparison of SLX and SDEM Coefficients}
\label{tab:slx-sdem}
\begin{tabular}{l d{2.4} d{2.4} d{2.4} c d{2.4} d{2.4} d{2.4} }
\toprule
 & \multicolumn{3}{c}{SLX} & & \multicolumn{3}{c}{SDEM} \\
\cmidrule{2-4}\cmidrule{6-8}
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.} & 
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.}\\
\midrule
```{r slx-sdem, results = "asis"}
slx_coefs <- lapply(slx_vs_sdem, function(x) tidy(x, conf.int = TRUE))
slx_stats <- lapply(slx_vs_sdem, glance)
tab <- right_join(slx_coefs$SLX, slx_coefs$SDEM, by = "term") %>%
  transmute(term, estimate.x, conf.low.x, conf.high.x, space = " ",
                  estimate.y, conf.low.y, conf.high.y) %>%
  mutate(term = unlist(cm))

xtab <- xtable(tab, digits = 4)
print(xtab, include.rownames = FALSE, include.colnames = F,
      sanitize.text.function = identity, floating = F, only.contents = T, hline.after = NULL)
```
\midrule
Log Likelihood & \multicolumn{3}{c}{$-4831.974$} & & \multicolumn{3}{c}{$-4435.507$}\\
Num.Obs.       & \multicolumn{3}{c}{$2099$}   & & \multicolumn{3}{c}{$2099$}\\
\bottomrule
\end{tabular}
\end{table*}

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and accessibility
coefficients for three models using different specifications of the utility
equation, as well as the ten-minute walk buffer calculated by ParkScore. 
Note that the model called "Size and Distance" is precisely the model
labeled "SDEM" in Table \ref{tab:slx-sdem}. The controlling covariates are
suppressed for clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in the Appendix. Each of the three
utility-based specifications is significant at the $\alpha = 0.05$ level or lower,
and is of the hypothesized directionality. That is, residents living in tracts 
with increased utility-based accessiblity to green spaces show a significantly
higher physical activity participation rate. Taking the rough mean of the 2
estimated utility-based coefficients, moving from the least-accessible tract
($A_i = -3$) to the
most-accessible tract ($A_i = 3$) is expected to raise the physical activity rate
by `r 6 * .15` percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor park access
is anywhere between `r 6 * .01` and `r 6 * .3` percentage points.

@@

\begin{table*}
\caption{Estimated Effect of Accessibility on Physical Activity Rates}
\label{tab:pa-models}
```{r pa-models, results = "asis"}
texreg(pa_models[c(2:4)], digits = 4, ci.force = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "lambda" = "$\\lambda$: spatial correlation", 
         "euc_access" = "Accessibility", 
         "euc_multi" = "Accessibility", 
         "walk10TRUE" = "Accessibility"),
       caption = "Estimated Effect of Accessibility on Physical Activity Rates",
       label = "tab:pa-models", table = FALSE,
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.  $y = $ tract-level physical activity rate.")
```
\end{table*}

@@

As far as the accessibility values are concerned, none of the three
utility-based specifications nor the ParkScore 10-minute walk buffer are
correlated with obesity rates at the traditional 95\% confidence interval. 
It is important to consider the implications of the full range of the confidence
interval, however [@Amrhein2019]. The results suggest that after controlling
for tract socioeconomic variables, physical activity rates, and spatially correlated
error terms, obesity rates in tracts with excellent park accessibility are expected 
to be between `r 6 * 0.113` percentage points lower and `r 6 * .0425` percentage 
points higher than tracts with poor park accessibility. This range of values 
suggests that accessibility to parks may have obesity-reducing effects beyond
the effect of accessibility on physical activity rates worth further study.

\begin{table*}
\caption{Estimated Effect of Accessibility on Obesity Rates}
\label{tab:obesity-models}
```{r obesity-models, results="asis"}
texreg(obesity_models[c(2:4)], digits = 4, ci.force = TRUE, 
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "physact" = "Physical Activity", 
         "lag.physact" = "$\\gamma$: Physical Activity",
         "lambda" = "$\\lambda$: spatial correlation", 
         "euc_access" = "Accessibility", 
         "euc_multi" = "Accessibility", 
         "walk10TRUE" = "Accessibility"),
       caption = "Estimated Effect of Accessibility on Obesity Rates",
       label = "tab:obesity-models", table = FALSE,
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.  $y = $ tract-level obesity rate.")
```
\end{table*}

@@

Finally, this study is primarily focused on the hypothesis that accessibility to
parks encourages physical activity, which in turn reduces obesity. There are a
multitude of other hypotheses that might be proposed and tested with the basic
methodology we have employed here, or competing explanations for the outcomes we
have observed. It is distinctly possible, for instance, that individuals who wish
to exercise regularly in parks choose to live near them. Given that the CDC models
informing the small-area obesity and physical activity estimates presumably include variables
likely to influence such preferences (income, etc.), our investigation cannot
isolate the preferences from the effect. Controlling for such a
self-selection effect would be necessary to isolate the exogenous
impacts of park access on obesity or other health outcomes. And regarding these
other variables: this study did not consider potential relationships between park
access and hospitalization rates, life expectancy, respiratory disease,
mental health, or any number of potential beneficial outcomes hypothesized
or explored in the existing literature. The COVID-19 pandemic and its associated
social distancing behaviors in many regions of the world has also drawn
attention to problems of park access, use, and equity. Exploring these
connections and their underlying mechanisms should be a priority as city
planners and urban architects attempt to improve the quality of life of urban
residents in the future.

@@

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
accessibility and these metrics, previous literature has produced mixed
findings, perhaps owing to the range of variables modeled and the coarse spatial
scale of the analyses. Using New York City as a case study, we presented a
holistic and flexible measurement for park accessibility that compares the
continuous distance to all parks in the region, weighted against the size of the
park and its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive relationship
where the least park-accessible tracts have an expected physical activity
participation rate between `r 6 * .01` and `r 6 * .3` percentage points lower than the most
accessible tracts. We also found a suggestive though not sigificant relationship
between park accessiblity and expected obesity rates *in addition to* the effect
of physical activity participation.
